Environmental pathogen reservoirs and habitat heterogeneity in a metapopulation

### 2014-03-17 Manuscript/manuscript.tex
Modelrun output and discussion edits
Clean up modelrun output to avoid loops and streamline.  Initial edits
of discussion and some plotting changes.

\date{May 30, 2013}

@@

This model framework demonstrates that environmental longevity increases pathogen persistence and spread.  This agrees both with our intuition -- environmental longevity should provide greater opportunity for transmission -- and with models of chronic wasting disease (\cite{Almberg2011},\cite{Sharp2011}).  \emph{Develop further -- both the papers cited, and other studies including a reservoir.  Remember to highlight not just consistency, but also novelty.}

When longevity is sufficient for the pathogen to persist, the variance in the patch quality distribution (i.e. habitat heterogeneity) controls how and where it spreads.  In particular, increased variance facilitates pandemic dynamics, increasing the likelihood that the pathogen will spread through the entire host population.  The mechanisms underlying this behavior can best be understood by drawing parallels between our metapopulation model and models of pathogen spread through a contact network.  In this framework, our metapopulation becomes a network with patches as nodes connected through immigration (with the strength of the connection measured by connectivity).  Many studies have shown that networks with high degree variance (where a node's degree is roughly equivalent to its connectivity in our metapopulation model) are easily invaded by pathogens, which are then able to spread much more rapidly than in less heterogeneous networks (\cite{Pastor-Satorras2001}).

>>>>

\date{November 26, 2013}

@@

The positive relationship between pathogen persistence/spread and environmental longevity demonstrated in the theoretical model above is well documented in models of empirical systems.  \cite{Almberg2011} find that, in plausible simulations of chronic wasting disease in deer, increasing prion survival in the environment increases the resulting force of infection and peak prevalence, with the  importance of reservoir transmission increasing through time.  Similarly, Breban et al 2009 show that environmental transmission of avian influenza virus allows it to persist in situations where direct transmission alone is insufficient.  \emph{Other examples.}

When longevity is sufficient for the pathogen to persist, we find that habitat heterogeneity controls how and where it spreads.  In particular, increased habitat quality variance facilitates pandemic dynamics, increasing the likelihood that the pathogen will spread through the entire host population.  The mechanisms underlying this behavior can best be understood by drawing parallels between our metapopulation model and models of pathogen spread through a contact network.  In this framework, our metapopulation becomes a network with patches as nodes connected through immigration (with the strength of the connection measured by connectivity).  Many studies have shown that networks with high degree variance (where a node's degree is roughly equivalent to its connectivity in our metapopulation model) are easily invaded by pathogens, which are then able to spread much more rapidly than in less heterogeneous networks (\cite{Pastor-Satorras2001}).
